Hydra Battles in Coq

### 2020-11-04 doc/hydras.tex
Proposition to rename the directory src/ into the more canonical theories/ (#7)

At any place of our development, you may use the  \texttt{Print Assumptions {\it ident}} command in order to verify on which axiom the theorem {\it ident} may depend. The following example is extracted from 
Library~\href{../src/html/hydras.Epsilon0.F_alpha.html}{hydras.Epsilon0.F\_alpha}, where we use the \texttt{coq-equations} plug-in (see Sect.~\vref{sect:wainer}).

@@

\subsection{Active Links}
The  links which appear in this pdf  document lead are of three possible kinds of destination:
\begin{itemize}
\item Local links to the document itself,
\item External links, mainly to \coq's page,
\item Local links to pages generated by \texttt{coqdoc}. According to the current makefile (through the commands \texttt{make html} and \texttt{make pdf}), 
  we assume that the page generated from a library \texttt{XXX/YYY.v} is stored as
the relative address \texttt{../src/html/hydras.XXX.YYY.html} (from the location of the pdf)
Thus,  active links towards our \coq{} modules may be incorrect if you got this \texttt{pdf} document otherwise than by compiling the distribution available in
\url{https://github.com/coq-community/hydra-battles}.

@@

Technically, a \emph{hydra} is just a finite ordered tree, each node of which 
has any number of sons. Note that, contrary to the computer science tradition, we will show the hydras 
with the heads up and the foot (i.e., the root of the tree) down.
Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}).

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}

@@

\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}

@@

\vspace{4mm}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}

@@

\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}

@@

\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}

@@

\vspace{4pt}
\noindent
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}

@@

\vspace{4pt}
\noindent
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}

@@

\vspace{4pt}\emph{From Module\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}

@@

\vspace{4pt}
\noindent
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}

@@

\vspace{4pt}\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
\label{types:Battle}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}

@@

% \vspace{4pt}
% \emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}

@@

The complete proof is in file \url{../src/html/hydras.Hydra.BigBattle.html}.

@@

We can make bigger steps through iterations of \texttt{next}.
The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
is defined and studied in~\href{../src/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.

@@

\begin{Coqsrc}
Definition exp2 n := iterate (fun n => 2 * n) n 1.
\end{Coqsrc}
Using some facts (proven in 
\href{../src/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}

@@

The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
\url{../src/html/hydras.Hydra.Hydra_Theorems.html} show that the classes \texttt{free} and \texttt{standard} satisfy this property.

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}

@@

Let us present the main steps of that proof, the script of which  is in the module ~\href{../src/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''. In effect, the elements of $\omega$, considered as a set, are just the natural numbers (see next chapter for more details)}.

@@

\vspace{4pt}
\noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}}

@@

\vspace{4pt}
\noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations/ON_Omega.html}{OrdinalNotations.ON\_Omega}}

@@

\vspace{4pt}
\noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations/ON_plus.html}{OrdinalNotations.ON\_plus}}

@@

\vspace{4pt}
\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}

@@

The following definitions are in Library \href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.

@@

\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}

@@

\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}}
\begin{Coqsrc}
  Definition ZeroLimitSucc_dec {A:Type}{lt: relation A}
           {compare : A -> A -> comparison}
           {on : ON lt compare} :=
  forall alpha,
    {Least alpha} +
    {Limit alpha} +
    {beta: A | Successor alpha beta}.
(* ... *)
\end{Coqsrc}

\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}

@@

\vspace{4pt}
\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Omega2.html}{ON\_Omega2}}

@@

The detailed  proof script is in the file \url{../src/html/hydras.Hydra.Omega2_Small.html}.

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}

@@

\vspace{4pt}
\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}

@@

\vspace{4pt}
\noindent\emph{From Module~\href{../src/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}

@@

\begin{remark}
It is important to keep in mind  that the integer $n$ is not an ``element'' of \texttt{FinOrd $n$}. In set-theoretic presentations of ordinals, the set associated with the ordinal $n$ is $\{0,1,\dots,n-1\}$. 
In our formalization, the interpretation of an ordinal as a set is realized by the following definition
(in ~\href{../src/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}).

@@

This exercise is partially solved for $i=3$ (in ~\href{../src/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).

@@

\begin{project}
The directory \texttt{src/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury. Every ordinal $\alpha$ is represented by a list $l$ whose elements are the coefficients of $\omega$ in  the Cantor normal form of $\alpha$ (in reverse order). For instance, the ordinal 
$\omega^{8}\times 5 + \omega^{6}\times 8 + \omega^2\times 10 + \omega + 7$ is represented by the list \texttt{[5;0;8;0;0.0;10,1,7]}.

@@

\begin{remark}
 The set of ordinal terms in Cantor normal form (see Chap.~\ref{chap:T1}) and 
in Veblen normal form (see 
\href{../src/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
\end{remark}

@@

\vspace{4pt}
\noindent\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#T1}{Epsilon0.T1}}

@@

\emph{Please look also at the library
\href{../src/html/hydras.Gamma0.T2.html}{Gamma0.T2.html}}.

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#compare}{Epsilon0.T1}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#compare_refl}{Epsilon0.T1}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#lt_irrefl}{Epsilon0.T1}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Prelude.Restriction.html}{Ordinals.Prelude.Restriction}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#LT}{Epsilon0.T1}}

@@

\emph{From Module~\href{../src/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}

@@

 \vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}

@@

Let \texttt{n:nat} and \texttt{beta:T1} such that \texttt{ocons alpha n beta} is in normal form. 
We prove first that \texttt{beta} is accessible,  which allows us to by well-founded induction on \texttt{beta}, 
and natural induction on \texttt{n}, that (\texttt{ocons alpha n beta}) is accessible.
The proof, quite long, can be consulted in \url{../src/html/hydras.Epsilon0.T1.html}

@@

\begin{remark}
\label{remark:a3pat}
The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the library \href{../src/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
 \end{remark}

@@

\emph{From Module~\href{../src/html/hydras.Schutte.Schutte.Correctness_E0.html}{Schutte.Schutte.Correctness\_E0}}

@@

\textbf{Hint}. You may use the following definition (in 
\href{../src/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}).

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}

@@

\item If $\beta<\alpha<\epsilon_0$, then there is a ``path'' from $\alpha$ to $\beta$, \emph{i.e.} a
sequence $\alpha_0=\alpha, \alpha_1, \dots, \alpha_n=\beta$, where for every $k<n$, there exists some $i_k$ such that $\alpha_{k+1}=\canonseq{\alpha_k}{i_k}$
\item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1})$ in one round with
the replication factor $i$ (Lemma \href{../src/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
\end{itemize}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.Canon.html\#canonS}{Epsilon0.Canon}}

@@

We did not  try to prove that our definition really implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
~\href{../src/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}

@@

\emph{From Module~\href{../src/html/hydras.Hydra.O2H.html}{Hydra.O2H}}

@@

The following lemma (proved in ~\href{../src/html/hydras.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.

@@

In the module \href{../src/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$. This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.

@@

\subsection{The case of Free Battles}
\label{sec:free-battles-case}
Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles). The following lemmas are proved in Module~\href{../src/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.

@@

\section{The Case of Standard Battles}
\label{sec:standard-intro}\label{std-case}
One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
Module~\href{../src/html/hydras.Hydra.O2H.html}{Hydra.O2H}.

@@

This lemma builds a battle out of any inequality $\beta<\alpha$. 
It is a straightforward application of \texttt{LT\_path\_to} of
Module~\href{../src/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}:

@@

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.KS.html}{Epsilon0.KS}}

@@

In the same way as for free battles, we import a large inequality 
from 
the module \href{../src/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}.

@@

On may ask whether the preceding results are not too restrictive, since they 
refer to a particular data type \texttt{T1}.
In fact, our representation of ordinals strictly less than 
 $\epsilon_0$ is faithful to their mathematical definition, at least 
Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
(please see also \href{../src/html/hydras.Schutte.Correctness_E0.html}{the module \texttt{Ordinals.Schutte.Correctness\_E0}}).

@@

\noindent
\emph{From Module~\href{../src/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}

>>>>

At any place of our development, you may use the  \texttt{Print Assumptions {\it ident}} command in order to verify on which axiom the theorem {\it ident} may depend. The following example is extracted from 
Library~\href{../theories/html/hydras.Epsilon0.F_alpha.html}{hydras.Epsilon0.F\_alpha}, where we use the \texttt{coq-equations} plug-in (see Sect.~\vref{sect:wainer}).

@@

\subsection{Active Links}
The  links which appear in this pdf  document lead are of three possible kinds of destination:
\begin{itemize}
\item Local links to the document itself,
\item External links, mainly to \coq's page,
\item Local links to pages generated by \texttt{coqdoc}. According to the current makefile (through the commands \texttt{make html} and \texttt{make pdf}), 
  we assume that the page generated from a library \texttt{XXX/YYY.v} is stored as
the relative address \texttt{../theories/html/hydras.XXX.YYY.html} (from the location of the pdf)
Thus,  active links towards our \coq{} modules may be incorrect if you got this \texttt{pdf} document otherwise than by compiling the distribution available in
\url{https://github.com/coq-community/hydra-battles}.

@@

Technically, a \emph{hydra} is just a finite ordered tree, each node of which 
has any number of sons. Note that, contrary to the computer science tradition, we will show the hydras 
with the heads up and the foot (i.e., the root of the tree) down.
Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}).

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hydra}{Hydra.Hydra\_Definitions}}

@@

\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#head}{Hydra.Hydra\_Definitions}}

@@

\vspace{4mm}
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html\#Hy}{Hydra.Hydra\_Examples}}

@@

\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#hcons_mult}{Hydra.Hydra\_Definitions}}

@@

\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}

@@

\vspace{4pt}
\noindent
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}

@@

\vspace{4pt}
\noindent
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}

@@

\vspace{4pt}\emph{From Module\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}

@@

\vspace{4pt}
\noindent
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#S1}{Hydra.Hydra\_Definitions}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#R2}{Hydra.Hydra\_Definitions}}

@@

\vspace{4pt}\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}}.

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#round_n}{Hydra.Hydra\_Definitions}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html}{Hydra.Hydra\_Definitions}}
\label{types:Battle}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#free}{Hydra.Hydra\_Definitions}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#standard}{Hydra.Hydra\_Definitions}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#fight}{Hydra.Hydra\_Definitions}}

@@

% \vspace{4pt}
% \emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html}{Hydra.Hydra\_Lemmas}}

@@

The complete proof is in file \url{../theories/html/hydras.Hydra.BigBattle.html}.

@@

We can make bigger steps through iterations of \texttt{next}.
The functional \texttt{iterate}, similar to Standard Library's \texttt{Nat.iter},
is defined and studied in~\href{../theories/html/hydras.Prelude.Iterates.html\#iterate}{Prelude.Iterates}.

@@

\begin{Coqsrc}
Definition exp2 n := iterate (fun n => 2 * n) n 1.
\end{Coqsrc}
Using some facts (proven in 
\href{../theories/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Alive}{Hydra.Hydra\_Definitions}}

@@

The theorems \texttt{Alive\_free} and \texttt{Alive\_standard} of the module 
\url{../theories/html/hydras.Hydra.Hydra_Theorems.html} show that the classes \texttt{free} and \texttt{standard} satisfy this property.

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Lemmas.html\#next_round_dec}{Hydra.Hydra\_Lemmas}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Definitions.html\#Hvariant}{Hydra.Hydra\_Definitions}}

@@

Let us present the main steps of that proof, the script of which  is in the module ~\href{../theories/html/hydras.Hydra.Omega_Small.html}{Hydra/Omega\_Small.v} \footnote{ The name of this file means ``the ordinal $\omega$ is too small for proving the termination of [free] hydra battles ''. In effect, the elements of $\omega$, considered as a set, are just the natural numbers (see next chapter for more details)}.

@@

\vspace{4pt}
\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}}

@@

\vspace{4pt}
\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations/ON_Omega.html}{OrdinalNotations.ON\_Omega}}

@@

\vspace{4pt}
\noindent\emph{From Library~\href{../theories/html/hydras.OrdinalNotations/ON_plus.html}{OrdinalNotations.ON\_plus}}

@@

\vspace{4pt}
\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}

@@

The following definitions are in Library \href{../theories/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.

@@

\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}

@@

\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}}
\begin{Coqsrc}
  Definition ZeroLimitSucc_dec {A:Type}{lt: relation A}
           {compare : A -> A -> comparison}
           {on : ON lt compare} :=
  forall alpha,
    {Least alpha} +
    {Limit alpha} +
    {beta: A | Successor alpha beta}.
(* ... *)
\end{Coqsrc}

\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega_plus_omega.html}{OrdinalNotations.ON\_Omega\_plus\_omega}}

@@

\vspace{4pt}
\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Omega2.html}{ON\_Omega2}}

@@

The detailed  proof script is in the file \url{../theories/html/hydras.Hydra.Omega2_Small.html}.

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}

@@

\vspace{4pt}
\emph{From Module ~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}

@@

\vspace{4pt}
\noindent\emph{From Module~\href{../theories/html/hydras.OrdinalNotations.ON_Finite.html}{OrdinalNotations.ON\_Finite}}

@@

\begin{remark}
It is important to keep in mind  that the integer $n$ is not an ``element'' of \texttt{FinOrd $n$}. In set-theoretic presentations of ordinals, the set associated with the ordinal $n$ is $\{0,1,\dots,n-1\}$. 
In our formalization, the interpretation of an ordinal as a set is realized by the following definition
(in ~\href{../theories/html/hydras.OrdinalNotations.Generic.html}{OrdinalNotations.Generic}).

@@

This exercise is partially solved for $i=3$ (in ~\href{../theories/html/hydras.OrdinalNotations.Example_3PlusOmega.html}{OrdinalNotations.Example\_3PlusOmega}).

@@

\begin{project}
The directory \texttt{theories/OmegaOmega} contains an ad-hoc formalization of $\omega^\omega$, contributed by Pascal Manoury. Every ordinal $\alpha$ is represented by a list $l$ whose elements are the coefficients of $\omega$ in  the Cantor normal form of $\alpha$ (in reverse order). For instance, the ordinal 
$\omega^{8}\times 5 + \omega^{6}\times 8 + \omega^2\times 10 + \omega + 7$ is represented by the list \texttt{[5;0;8;0;0.0;10,1,7]}.

@@

\begin{remark}
 The set of ordinal terms in Cantor normal form (see Chap.~\ref{chap:T1}) and 
in Veblen normal form (see 
\href{../theories/html/hydras.Gamma0.Gamma0.html}{Gamma0.Gamma0}) are shown to be ordinal notation systems, but there is a lot of work to be done in order to unify ad-hoc  definitions and proofs which were written before the definition of the \texttt{ON} type class.
\end{remark}

@@

\vspace{4pt}
\noindent\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#T1}{Epsilon0.T1}}

@@

\emph{Please look also at the library
\href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2.html}}.

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare}{Epsilon0.T1}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#compare_refl}{Epsilon0.T1}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#lt_irrefl}{Epsilon0.T1}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#nf_b}{Epsilon0.T1}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Prelude.Restriction.html}{Ordinals.Prelude.Restriction}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#LT}{Epsilon0.T1}}

@@

\emph{From Module~\href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}}

@@

 \vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.T1.html\#succ}{Epsilon0.T1}}

@@

Let \texttt{n:nat} and \texttt{beta:T1} such that \texttt{ocons alpha n beta} is in normal form. 
We prove first that \texttt{beta} is accessible,  which allows us to by well-founded induction on \texttt{beta}, 
and natural induction on \texttt{n}, that (\texttt{ocons alpha n beta}) is accessible.
The proof, quite long, can be consulted in \url{../theories/html/hydras.Epsilon0.T1.html}

@@

\begin{remark}
\label{remark:a3pat}
The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering~\cite{DershowitzRPO, a3pat} is available in the library \href{../theories/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
 \end{remark}

@@

\emph{From Module~\href{../theories/html/hydras.Schutte.Schutte.Correctness_E0.html}{Schutte.Schutte.Correctness\_E0}}

@@

\textbf{Hint}. You may use the following definition (in 
\href{../theories/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}).

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Termination.html\#m}{Hydra.Hydra\_Termination}}

@@

\item If $\beta<\alpha<\epsilon_0$, then there is a ``path'' from $\alpha$ to $\beta$, \emph{i.e.} a
sequence $\alpha_0=\alpha, \alpha_1, \dots, \alpha_n=\beta$, where for every $k<n$, there exists some $i_k$ such that $\alpha_{k+1}=\canonseq{\alpha_k}{i_k}$
\item Canonical sequences correspond tightly to rounds of hydra battles: if $\alpha\not=0$,
then $\iota(\alpha)$ is transformed into $\iota(\canonseq{\alpha}{i+1})$ in one round with
the replication factor $i$ (Lemma \href{../theories/html/hydras.Hydra.O2H.html\#canonS_iota_i}{Hydra.O2H.canonS\_iota\_i}).
\item From the two previous properties, we infer that whenever $\beta<\alpha<\epsilon_0$, there exists a (free) battle from $\iota(\alpha)$ to $\iota(\beta)$.
\end{itemize}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS}{Epsilon0.Canon}}

@@

We did not  try to prove that our definition really implements Ketonen and Solovay's  \cite{KS81}'s canonical sequences. The most important is that we were able to prove the 
abstract properties  of canonical sequences that are really used in our proof. The complete proofs are in the module
~\href{../theories/html/hydras.Epsilon0.Canon.html}{Epsilon0.Canon}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}

@@

\emph{From Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}}

@@

The following lemma (proved in ~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H.v}) maps  canonical sequences to rounds of hydra battles.

@@

In the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}, we assume there exists some variant $m$ bounded by some ordinal $\mu<\epsilon_0$. This part of the development is parameterized by some class $B$ of battles, which will be instantiated later to \texttt{free} or \texttt{standard}.

@@

\subsection{The case of Free Battles}
\label{sec:free-battles-case}
Let us now consider that $B$ is instantiated to \texttt{free} (which means that we are considering proofs of termination of \emph{all} battles). The following lemmas are proved in Module~\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Free.html}{Hydra.Epsilon0\_Needed\_Free}.
The case $B=\texttt{standard}$ is studied in section~\vref{std-case}.

@@

\section{The Case of Standard Battles}
\label{sec:standard-intro}\label{std-case}
One may wonder if our theorem holds also in the framework of standard battles. Unfortunately, its proof relies on the lemma \texttt{LT\_to\_round\_plus} of
Module~\href{../theories/html/hydras.Hydra.O2H.html}{Hydra.O2H}.

@@

This lemma builds a battle out of any inequality $\beta<\alpha$. 
It is a straightforward application of \texttt{LT\_path\_to} of
Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}:

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.KS.html}{Epsilon0.KS}}

@@

In the same way as for free battles, we import a large inequality 
from 
the module \href{../theories/html/hydras.Hydra.Epsilon0_Needed_Generic.html}{Hydra.Epsilon0\_Needed\_Generic}.

@@

On may ask whether the preceding results are not too restrictive, since they 
refer to a particular data type \texttt{T1}.
In fact, our representation of ordinals strictly less than 
 $\epsilon_0$ is faithful to their mathematical definition, at least 
Kurt Schütte's~\cite{schutte}, as proved in Chapter~\vref{chap:schutte}.
(please see also \href{../theories/html/hydras.Schutte.Correctness_E0.html}{the module \texttt{Ordinals.Schutte.Correctness\_E0}}).

@@

\noindent
\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}
